Interest Flooding Attack and Countermeasures in Named Data Networking

### 2013-01-15 05-evaluation.tex
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@936 df1a90cd-247b-4364-bdaf-287a704eb499

The metric we choose to quantify the effectiveness of our algorithms is the {\it percentage of satisfied Interests for legitimate users}. This metric corresponds to the quality of service experienced by legitimate users when the network is under attack. In other words, if the network implements a mitigation method $X$ and a high percentage of user-expressed Interests are satisfied even while the network is under attack, then one can conclude that method $X$ is highly effective at mitigating the attack. We also ensure that all Interests expressed by legitimate users during a period of no-attack are satisfied.

@@

% Traffic pattern
In our experiments, we assume that legitimate users express Interests at constant average rates with randomized time gap between two consecutive Interests, where the random number for the gap follows a uniform distribution. We believe that this traffic pattern provides a reasonable approximation of traffic mix from all network users without excessive buffering. To quantify the behavior of our mitigation strategies under a worse case attack scenario, we ensure that all the attackers send junk Interests as fast as they can. 
%(remember, that Interest limits will not allow real flooding of Intersests in all of the designed mitigation algorithms).

@@

% Fixed parameters
We set the \emph{delay} and \emph{data size} parameters for the Interest limit calculation to a fixed value for every node in the simulated topology. In particular, for the small-scale binary tree topology, we set delay to 80~ms, while for the large-scale ISP topology we set it to 330~ms (the order of the largest RTT). The data size is 1100~bytes for all simulation runs and topologies.

@@

We observe a successful DDoS attack, where 40\% attackers succeed in significantly shutting down the remaining 60\% legitimate users---a mere 15\% of their Interests are satisfied by Data from the producer. Contrary to expectations, the 60\% legitimate users do not receive at least 60\% of network resources. As described in the previous section, the key limitation of this algorithm is that it still admits Interests from attackers, a good percentage of which traverse all the way to the data producer. Significantly reducing the number of accepted malicious Interests is fundamental for improving the effectiveness of any mitigation algorithm.

@@

The effectiveness of this algorithm stems from the fact that routers do not admit malicious Interests into the network, thereby ensuring availability of network resources for serving legitimate users. The observed periodic dips in the Interest satisfaction ratios of legitimate users in Fig.~\ref{fig:small-scale attack progress} is a direct result of Interest satisfaction rate statistics decaying with time. The 50-second period approximately corresponds to the selected exponential decaying parameter $\alpha=e^{−1.0/30.0}$, which decays statistics to $1/e$ of the initial value that ranges from 30~seconds to 50~seconds. 
%The primary reason that such minimum peaks exist is the fact that 
When Interests from attackers start to get readmitted, they cause degradation of statistics on routers close to the producer (i.e., routers that observe traffic mix from legitimate and malicious users). Consequently, this degradation reduces the probability of legitimate Interests getting through (see Section~\ref{sec:probabilistic}) until malicious Interests are ``pushed back'' to the edge.

@@

This mitigation algorithm is able to effectively shut down attackers and ensure that almost all of the Interests from legitimate users are satisfied. 
%The only potential problem with the satisfaction-based pushback algorithm is that it features 
We observe a  sharp dip in the satisfaction ratio curve at the start of the attack.  It takes a few seconds for all routers to be fully aware of the attack, which happens when malicious Interests start to time out and explicit Interest limit announcements succeed in containing malicious Interests close to the attacker. Till then, the network for a short period of time (under 10~seconds for all simulation runs), fails to provide 100\% service for legitimate users. Once the malicious Interests are effectively shut down, all Interests from legitimate users are satisfied. Unlike the satisfaction-based Interest acceptance scenario, we do not observe any periodic dips in the satisfaction curve, as the pushback algorithm effectively guarantees that once an Interest is admitted, it will almost certainly be routed to the data producer.

\subsubsection{Network reaction with to varying number of attackers}

Our next goal is to study the effectiveness of our mitigation algorithms as a function of increasing adversaries in the network.
To this end, we vary the percentage of attackers in the topology from 6\% to over 50\%. Since the total number of end users in the topology is constant, as the number of attackers increases, the number of legitimate users decreases. All other parameters and experimental set-up are similar to the previous experiment. As before, for each mitigation algorithm, we perform 10 independent simulation runs.
%The second set of conducted experiments was aimed to answer the question of the effect and quality of the Interest flooding attack mitigation algorithm under different attack volume.
%To do this, for each algorithm we varied the number of adversary nodes in the topology, keeping the total number of client nodes constant: 1 attacker and 15 legitimate users, 3 attackers and 13 legitimate users, etc.

@@

In Fig.\ref{fig:small-scale-topo boxplot} we present  the Interest satisfaction ratio for legitimate users aggregated over the 10 simulation runs for the first minute of the attack. The results are as expected---for all three mitigation algorithms, as the percentage of attackers in the network increases, the lower is the Interest satisfaction ratio for legitimate users.
In the case of token bucket with per-interface fairness algorithm, just 3 attackers can halve the quality of service for the remaining 13 legitimate users. While both intelligent attack mitigation algorithms also show a decline in service quality as the percentage of attackers increases, this decline is much more gradual and marginal. In the case of satisfaction-based pushback algorithm, over 90\% of Interests from legitimate users are satisfied even when over 50\% of the users are malicious.

@@

Next, our goal is to understand the behavior of our mitigation strategies when deployed on a large-scale real network topology. Our ISP topology is based on a modified version of Rocketfuel's AT\&T topology~\cite{rocketfuel}.
%In order to approximate the general structure of the Internet,
% (scale-free structure, customer-provider, and peer-to-peer relations)
We extract the largest connected component comprising of 562 nodes from this original topology and separate the nodes into three categories: clients, gateways, and backbones. Nodes having degree less than four are classified as clients (344 red nodes as shown in Fig.~\ref{fig:large-scale}), nodes directly connected to clients are classified as gateways (109 green nodes), and the remaining nodes as classified as backbones (109 blue nodes). 
%(To ensure that paths in the topology are ``valley-free,'' we augmented the topology with necessary backbone-to-backbone links.) 
We assign bandwidth and  delay values to links based on their type---both values are random numbers within the respective ranges as shown in Table~\ref{tab:large-scale}. We experiment with placing the data producer at both a gateway node as well as backbone node, which we randomly pick for each simulation run. Similar to the binary tree topology experiments, we fix the number of malicious nodes at approximately 40\% (140 out of 344 client nodes in the topology) and randomly pick these nodes for each simulation run. We conduct 10 simulation runs for each mitigation algorithm, with the attack duration spanning a 5-minute interval.

@@

In Fig.~\ref{fig:large-scale}, we summarize our results aggregated over 10 simulations runs for each mitigation algorithm for the scenario, where the data producer is placed at a gateway node. We observe similar results for the data producer placed at the backbone node as well. As in the case of the binary-tree topology experiments, token bucket with per interface fairness is the most ineffective algorithm and satisfaction-based pushback is the most effective one.   Interest satisfaction percentage for legitimate users are approximately 25\% and almost 100\% respectively for these two mitigation methodologies.

@@

Satisfaction-based Interest acceptance algorithm, which was very effective for binary-tree topology, is completely ineffective when deployed in a larger realistic topology. For the duration of the attack, legitimate users experience poor quality of service with only 25\% of their Interests being satisfied and continue to experience degraded service long after the attack has stopped. This poor performance, as detailed in  in Section~\ref{sec:probabilistic}, is due to the fact that each router in the path makes an independent, uncoordinated decision on whether to forward or drop an Interest. In the case of the larger-topology, with much higher average hop count, Interest packets from legitimate users have a very low probability of reaching the data producer, resulting in poor Interest satisfaction statistics, which in turn leads to further penalization of new Interests from them.

To summarize, in all of our simulations, the satisfaction-based pushback algorithm is the most effective technique as it restricts malicious Interests from even entering the network. 
%The only short periods of time when malicious Interests are getting admitted to the network is when routers have either no prior knowledge about per-interface satisfaction ratios (the initial period of the attack) or such knowledge becomes stale (statistics decaying during the attack). As soon as the knowledge is obtained or refreshed, the service for legitimate users returns to norm.

@@

This paper is a first step in understanding the impact of Interest flooding attacks in NDN and exploring the solution space. 
While designing our mitigation algorithms, we exploit two key features of NDN architecture, namely routers maintaining state about the Interests they have forwarded and Data traffic taking the reverse path of the Interest traffic. 
We test the efficacy of our algorithms by simulating them on both a binary-tree topology designed to test the worst-case effectiveness of our algorithms as well as on a realizing ISP-based topology designed to provide insights of our algorithms' behavior when deployed on a real network. 
While our results are promising and show a great potential in mitigating Interest flooding DDoS attacks, there are certain limitations which should be addressed in future research.

First, in our evaluations we used a simple and static attacker model---attackers send junk Interests as fast as possible. 
In the future work, we plan to explore the impact of models where the attackers are more sophisticated and dynamically adapt their behavior and junk Interest sending patterns based on the network reaction. 
Second, we made an assumption that Interests are not satisfied by an intermediate router's cache and always forwarded all the way to the producer.  
The future work needs to study impact of Interest flooding attack in more realistic scenarios, including investigation effects of multi-path routing and presence of in network caches.
% We also ignored NDN's strategy layer that routes around failure and congestion in the network. 
% In future work, we plan to study the impact of turning on these features. 
Finally, we also plan to perform more extensive simulations on other realistic Internet-like topologies and traffic patterns.

>>>>

The metric we choose to quantify the effectiveness of our algorithms is the {\it percentage of satisfied Interests for legitimate users}. This metric corresponds to the quality of service experienced by legitimate users when the network is under attack. In other words, if the network implements a mitigation method $X$ and a high percentage of user-expressed Interests are satisfied even while the network is under attack, then one can conclude that method $X$ is highly effective at mitigating the attack.% We also ensure that all Interests expressed by legitimate users during a period of no-attack are satisfied.

@@

% Traffic pattern
In our experiments, we assume that legitimate users express Interests at constant average rates with randomized time gap between two consecutive Interests, where the random number for the gap follows a uniform distribution. We believe that this traffic pattern provides a reasonable approximation of traffic mix from all network users without excessive buffering. To quantify the behavior of our mitigation strategies under a worse case attack scenario, we assume all the attackers send junk Interests as fast as they can, no Interest --including legitimate ones-- can be satisfied from caches, routers are configured for single-path routing and there is only one, single-homed producer for the prefix under attack.
%(remember, that Interest limits will not allow real flooding of Intersests in all of the designed mitigation algorithms).

@@

% Fixed parameters
We set the \emph{delay} and \emph{data size} parameters for the Interest limit calculation to an initial %fixed 
value for every node in the simulated topology. In particular, for the small-scale binary tree topology, we set delay to 80~ms, while for the large-scale ISP topology we set it to 330~ms (the order of the largest RTT). The data size is 1100~bytes for all simulation runs and topologies.
% THIS IS BULLSHIT! There is absolutely no logic to set the delay to the same value on all nodes in the network. We discussed multiple times that you should fix it so that the delays are realistic and gets smaller as the nodes gets closer to the producer...

@@

We observe a successful DDoS attack, where 40\% attackers succeed in significantly shutting down the remaining 60\% legitimate users---a mere 15\% of their Interests are satisfied by Data from the producer. Contrary to expectations, the 60\% legitimate users do not receive at least 60\% of network resources. As described in the previous section, the key limitation of this algorithm is that it still admits Interests from attackers and causes congestion and packet drops near the victim. %Significantly reducing the number of accepted malicious Interests is fundamental for improving the effectiveness of any mitigation algorithm.

@@

The effectiveness of this algorithm stems from the fact that routers can differentiate and limit malicious Interests into the network%, thereby ensuring availability of network resources for serving legitimate users
. The observed periodic dips in the Interest satisfaction ratios of legitimate users in Fig.~\ref{fig:small-scale attack progress} is a direct result of Interest satisfaction rate statistics decaying with time. The 50-second period approximately corresponds to the selected exponential decaying parameter $\alpha=e^{−1.0/30.0}$, which decays statistics to $1/e$ of the initial value that ranges from 30~seconds to 50~seconds. 
%The primary reason that such minimum peaks exist is the fact that 
When Interests from attackers start to get readmitted, they cause degradation of statistics on routers close to the producer (i.e., routers that observe mixed traffic from legitimate and malicious users). Consequently, this degradation reduces the probability of legitimate Interests getting through (see Section~\ref{sec:probabilistic}) until malicious Interests are ``pushed back'' to the edge.

@@

In our simulations, this mitigation algorithm was able to effectively shut down attackers and ensure almost all the Interests from legitimate users are satisfied. 
%The only potential problem with the satisfaction-based pushback algorithm is that it features 
We observe a  sharp dip in the satisfaction ratio curve at the start of the attack as it takes a few seconds for all routers to be fully aware of the attack. Recovery quickly happens as malicious Interests start to time out and explicit Interest limit announcements start to succeed in containing malicious Interests close to the attacker. Till then, the network, for a short period of time (under 10~seconds for all simulation runs), fails to provide 100\% service for legitimate users. Once the malicious Interests are effectively shut down, all Interests from legitimate users are satisfied. Unlike the satisfaction-based Interest acceptance scenario, we do not observe any periodic dips in the satisfaction curve, as the pushback algorithm effectively guarantees that once an Interest is admitted, it will likely be routed all the way to the data producer.

\subsubsection{Network reaction to varying number of attackers}

Our next goal is to study the effectiveness of our mitigation algorithms as a function of increasing adversaries in the network.
To this end, we vary the percentage of attackers in the topology from 6\% to over 50\%. Since the total number of end users in the topology is constant, as the number of attackers increases, the number of legitimate users decreases. All other parameters and experimental set-up is the same with the previous experiment. As before, for each mitigation algorithm, we perform 10 independent simulation runs.
%The second set of conducted experiments was aimed to answer the question of the effect and quality of the Interest flooding attack mitigation algorithm under different attack volume.
%To do this, for each algorithm we varied the number of adversary nodes in the topology, keeping the total number of client nodes constant: 1 attacker and 15 legitimate users, 3 attackers and 13 legitimate users, etc.

@@

In Fig.\ref{fig:small-scale-topo boxplot} we present  the Interest satisfaction ratio for legitimate users aggregated over the 10 simulation runs for the first minute of the attack. The results are as expected---for all three mitigation algorithms, as the percentage of attackers in the network increases, the lower is the Interest satisfaction ratio for legitimate users.
In the case of token bucket with per-interface fairness algorithm, just 3 attackers can halve the quality of service for the remaining 13 legitimate users. While the two intelligent attack mitigation algorithms also show a decline in service quality as the percentage of attackers increases, this decline is much more gradual and marginal. In the case of satisfaction-based pushback algorithm, over 90\% of Interests from legitimate users are satisfied even if the 50\% of end nodes are malicious.

@@

In this section, we investigate the behavior of our mitigation strategies under a more realistic, large-scale network topology. The ISP topology we employed is based on a modified version of Rocketfuel's AT\&T topology~\cite{rocketfuel}.
%In order to approximate the general structure of the Internet,
% (scale-free structure, customer-provider, and peer-to-peer relations)
We extract the largest connected component comprising of 562 nodes from this original topology and separate the nodes into three categories: clients, gateways, and backbones. Nodes having degree less than four are classified as clients (344 red nodes as shown in Fig.~\ref{fig:large-scale}), nodes directly connected to clients are classified as gateways (109 green nodes), and the remaining nodes as classified as backbones (109 blue nodes). 
%(To ensure that paths in the topology are ``valley-free,'' we augmented the topology with necessary backbone-to-backbone links.) 
We assign bandwidth and  delay values to links based on their type---both values are random numbers within the respective ranges as shown in Table~\ref{tab:large-scale}. We experiment with placing the data producer at both a gateway node as well as backbone node, which we randomly pick for each simulation run. Similar to the binary tree topology experiments, we fix the number of malicious nodes at approximately 40\% (140 out of 344 client nodes in the topology) and randomly pick these nodes for each simulation run. We conduct 10 simulation runs for each mitigation algorithm, with the attack duration spanning a 5-minute interval.

@@

In Fig.~\ref{fig:large-scale}, we summarize our results aggregated over 10 simulations runs for each mitigation algorithm for the scenario, where the data producer is placed at a gateway node. We observe similar results for the data producer placed at the backbone node as well. As in the case of the binary-tree topology experiments, we observe that the token bucket with per interface fairness is the most ineffective algorithm and satisfaction-based pushback is the most effective one.  Interest satisfaction percentage for legitimate users are approximately 25\% and close to 100\% respectively for these two mitigation methodologies.

@@

Satisfaction-based Interest acceptance algorithm, which was very effective for binary-tree topology, is completely ineffective when deployed in a larger and more realistic topology. For the duration of the attack, legitimate users experience poor quality of service with only 25\% of their Interests being satisfied and continue to experience degraded service long after the attack has stopped. This poor performance, as detailed in Section~\ref{sec:probabilistic}, is due to the fact that each router in the path makes an independent, uncoordinated decision on whether to forward or drop an Interest. In the case of the larger-topology, with much higher average hop count, Interest packets from legitimate users have a very low probability of reaching the data producer, resulting in poor Interest satisfaction statistics, which in turn leads to further penalization of new Interests from them.

All our evaluations leaves us with the conclusion that, among the three techniques we tested under various topology and attacker concentrations, the satisfaction-based pushback is the most effective one in mitigating interest flooding attacks.% and it restricts malicious Interests from even entering the network. 
%The only short periods of time when malicious Interests are getting admitted to the network is when routers have either no prior knowledge about per-interface satisfaction ratios (the initial period of the attack) or such knowledge becomes stale (statistics decaying during the attack). As soon as the knowledge is obtained or refreshed, the service for legitimate users returns to norm.

@@

This paper is a first step in understanding the impact of Interest flooding attacks in NDN and exploring the solution space. 
While designing our mitigation algorithms, we exploit two key features of NDN architecture, namely routers maintaining state about the Interests they have forwarded and Data traffic taking the reverse path of the Interest traffic. 
We test the efficacy of our algorithms by simulating them on both a simple binary-tree topology % designed to test the worst-case effectiveness of our algorithms as well as on a realizing 
as well as a more realistic ISP-based topology designed to provide insights of our algorithms' behavior when deployed on a real network. 
While our results are promising and show a great potential in mitigating Interest flooding DDoS attacks, there are certain limitations which should be addressed in future research.

First, in our evaluations we used a simple and static attacker model---attackers send junk Interests as fast as possible. 
In the future work, we plan to explore the impact of models where the attackers are more sophisticated and dynamically adapt their behavior and Interest sending patterns based on the network reaction. 
Second, we made an assumption that Interests are not satisfied by an intermediate router's cache and always forwarded all the way to the producer.  
The future work needs to study impact of Interest flooding attack in more realistic scenarios, including investigation effects of multi-path routing, more realistic traffic patterns, and presence of in network caches.
% We also ignored NDN's strategy layer that routes around failure and congestion in the network. 
% In future work, we plan to study the impact of turning on these features. 
%Finally, we also plan to perform more extensive simulations on other realistic Internet-like topologies and traffic patterns.
